R-LIME: LIME法における矩形制約と最適化

### 2024-03-04 paper/paper.tex
Sort out repository

\def\mode{0} % 0: Graduation thesis, 1: ICPR paper

\ifnum\mode=0
	\documentclass[11pt]{article}
	\usepackage{luatexja}
	\usepackage{prmlthesis}
	\usepackage[a4paper]{geometry}
\else
	\documentclass[runningheads]{llncs}
\fi

@@

% import the TeX file containing macro definitions
\input{macro}

@@

\begin{figure}[tbp]
	\def\scale{0.38}
	\centering
	\vspace{0.5cm}
	\begin{subfigure}[t]{0.55\textwidth}
		\centering
		\includegraphics[scale=\scale]{example-instance}
		\caption{Focal point.}\label{fig:example-instance}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.45\textwidth}
		\begin{subfigure}[t]{\textwidth}
			\centering
			\includegraphics[scale=\scale]{example-lime}
			\caption{%
				Explanation by LIME\@.
				It provides contributions of each feature to the output,
				but does not explicitly indicate its scope.
			}\label{fig:example-lime}
			\vspace{0.4cm}
		\end{subfigure}
		\begin{subfigure}[t]{\textwidth}
			\centering
			\includegraphics[scale=\scale]{example-anchor}
			\caption{%
				Explanation by Anchor.
				It provides the effective scope of the explanation,
				but does not show the influence of each feature.
			}\label{fig:example-anchor}
		\end{subfigure}
	\end{subfigure}
	\hspace{0.3cm}
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\ifnum\mode=0
			\vspace{-1.79cm}
		\else
			\vspace{-2.24cm}
		\fi
		\includegraphics[scale=\scale]{example-rlime}
		\caption{%
			Explanation by R-LIME\@.
			It provides both contributions of each feature and
			its effective scope.
		}\label{fig:example-rlime}
	\end{subfigure}
	\caption[Example of explanations by LIME, Anchor and R-LIME]{%
		Example of explanations by LIME~\cite{ribeiro2016why},
		Anchor~\cite{ribeiro2018anchors} and R-LIME (our proposed method).
		for a sentiment prediction model.
	}\label{fig:example}
\end{figure}
In this paper,
we focus on local model-agnostic methods.
An example of explanations by LIME and Anchor
for a sentiment prediction model is illustrated in \cref{fig:example}.
LIME~\cite{ribeiro2016why} approximates the complex decision boundary
locally using a linear model around the given focal point,
then provides the weights of the linear model as the contribution of each feature
to the output.
The explanation by LIME (\cref{fig:example-lime}) provides suggests that
the word ``not'' mainly contributes to the positive prediction,
but does not explicitly indicate the effective scope.
Without the scope,
users might mistakenly apply the knowledge derived from the explanation
to other instances far from the focal point,
potentially leading to misunderstanding of the black-box model's behavior
\cite{ribeiro2018anchors}.
For example,
user may apply the insights derived from the explanation
to the sentence ``This book is not good.''
and mistakenly conclude that the word ``not''
mainly contributes to the positive prediction for this sentence as well,
which is incorrect obviously.
Anchor~\cite{ribeiro2018anchors} represents a rectangular region
containing the focal point,
that maximizes the probability of the black-box classifier outputting
the same label as the focal point.
While Anchor provides an effective scope of the explanation,
the information it includes is less than LIME,
limiting its utility~\cite{ribeiro2018anchors}.
The explanation by Anchor (\cref{fig:example-anchor})
suggests that changing words other than ``not'' and ``bad''
has little impact on the classifier's output.
While it clearly cannot be applied to the sentence ``This book is not good''
because of not including the word ``bad'',
the explanation does not provide details about the influence of each word,
resulting in less user insight into the model's behavior compared to LIME.

@@

\subsection{Algorithm}\label{sec:alg}
{%
	\begin{figure}[t]
		\centering
		\includegraphics[width=0.32\textwidth]{visual-rlime1}
		\includegraphics[width=0.32\textwidth]{visual-rlime2}
		\includegraphics[width=0.32\textwidth]{visual-rlime3}
		\caption[Overview of the R-LIME algorithm]{%
			Overview of the R-LIME algorithm.
			The progression of the algorithm is illustrated from left to right.
			The solid line represents the rectangular region $A$,
			and the dashed line represents the linear approximation model $g$
			learned within $A$.
			The initial value of $A$ is an empty rule (entire input space),
			and predicates are added to $A$, reducing coverage.
			The process continues until $\Prec(A)\ge\tau$,
			at which point the rule with the maximum coverage is output.
		}
	\end{figure}
	\def\myidt{\hspace{\algorithmicindent}}
	\ifnum\mode=0
		\begin{algorithm}[t]
			\small
			\input{alg1}
		\end{algorithm}
	\else
		\begin{algorithm}[p]
			\small
			\input{alg1}
		\end{algorithm}
	\fi

	\foreach \alg in {2,3,4}{%
			\begin{algorithm}[p]
				\small
				\input{alg\alg}
			\end{algorithm}
		}
}

@@

		{%
			\def\AB#1{\ifnum#1=0 A\else B\fi}
			\def\mylabel#1{\ifnum#1=0 \label{fig:A-instance}\else \label{fig:B-instance}\fi}
			\ifnum\mode=0
				\renewcommand{\arraystretch}{0.97}
			\else
				\renewcommand{\arraystretch}{1.02}
			\fi
			\begin{figure}[tbp]
				\foreach\a in {0,1}{%
						\centering
						\begin{subfigure}{\textwidth}
							\ifnum\mode=0
								\small
							\fi
							\centering
							\begin{tabular}{p{14em}m{16em}}
								\toprule
								\csvreader[no head, late after line= \\]{%
									\dir/\index{\a}.csv
								}{}{%
								\ifnum\thecsvrow=16 \midrule\fi\csvcoli & \csvcolii
								}
								\bottomrule
							\end{tabular}
							\caption{Instance~\AB{\a}}\mylabel{\a}
							\vspace{15pt}
						\end{subfigure}
					}
				\vspace{-15pt}
				\caption[Two instances sampled from recidivism dataset]{%
					Two instances sampled from recidivism dataset.
				}\label{fig:instance}
			\end{figure}
		}
		{%
			\ifnum\mode=0
				\def\scale{0.32}
				\def\imgwidth{0.45\textwidth}
				\def\hspacebase{\hspace{-0.5em}}
			\else
				\def\scale{0.315}
				\def\imgwidth{0.495\textwidth}
				\def\hspacebase{\hspace{-1.5em}}
			\fi
			\def\vspacebase{\vspace{0.5em}}
			\def\vspacebeforecaption{\vspace{-0.4em}}
			\begin{figure}[p]
				\centering
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{1.0em}
					\includegraphics[scale=\scale]{\dir/lime-\Asample}
					\vspacebeforecaption
					\caption{LIME}\label{fig:A-lime}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{0.5em}
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-70}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.70$)}\label{fig:A-rlime-70}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-80}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.80$)}\label{fig:A-rlime-80}
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{-0.5em}
					\includegraphics[scale=\scale]{\dir/newlime-\Asample-90}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.90$)}\label{fig:A-rlime-90}
				\end{subfigure}
				\caption[Explanation for Instance A by LIME and R-LIME]{%
					Explanation for Instance A by LIME and R-LIME.
				}\label{fig:A}
			\end{figure}
			\begin{figure}[p]
				\centering
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{-0.7em}
					\includegraphics[scale=\scale]{\dir/lime-\Bsample}
					\vspacebeforecaption
					\caption{LIME}\label{fig:B-lime}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{0.5em}
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-70}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.70$)}\label{fig:B-rlime-70}
					\vspacebase
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{-0.8em}
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-80}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.80$)}\label{fig:B-rlime-80}
				\end{subfigure}
				\begin{subfigure}[t]{\imgwidth}
					\hspacebase
					\hspace{-0.8em}
					\includegraphics[scale=\scale]{\dir/newlime-\Bsample-90}
					\vspacebeforecaption
					\caption{R-LIME ($\tau=0.90$)}\label{fig:B-rlime-90}
				\end{subfigure}
				\caption[Explanation for Instance B by LIME and R-LIME]{%
					Explanation for Instance B by LIME and R-LIME.
				}\label{fig:B}
			\end{figure}
		}
	}
\subsubsection{Experimental Results}
The results of the experiment are shown in \cref{fig:A,fig:B}.
The values assigned to each attribute name represent the contribution
(weights of the learned linear classifier)
to the output of the black-box classifier,
normalized such that the absolute sum is 1.
The figures also display the 5 attributes with the highest absolute contribution.

@@

\subsection{Changes in Reward Distribution in Best Arm Identification}\label{sec:reward}
{%
	\renewcommand{\arraystretch}{1.1}
	\begin{table}[tbp]
		\centering
		\caption[Deviation between the estimated accuracy and the true accuracy]{%
			Deviation between the estimated accuracy and the true accuracy.
			Deviation was relatively small considering confidence level $1-\delta=0.95$.
		}\label{tab:reward}
		\input{tab_exp_adv}
	\end{table}
}
In R-LIME,
the problem of selecting the rule with the highest accuracy is formulated
as the optimal arm identification problem in multi-armed bandit theory,
solved using the KL-LUCB algorithm \cite{kaufmann2013information}.
However, this algorithm assumes that the reward distribution remains constant,
while in R-LIME,
the reward distribution (accuracy of the linear approximation)
changes with every update of the approximation model after sampling.
Therefore, rewards obtained at an early stage
might influence the estimated value and deviate from the true value.

@@

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\ifnum\mode=0
	\bibliographystyle{prml}
\else
	\bibliographystyle{splncs04}
\fi
\bibliography{MyRef/ref}
%
\end{document}

>>>>

\def\mode{1} % 0: Graduation thesis, 1: ICPR paper

\ifnum\mode=0
  \documentclass[11pt]{article}
  \usepackage{luatexja}
  \usepackage{style/prmlthesis}
  \usepackage[a4paper]{geometry}
\else
  \documentclass[runningheads]{llncs}
\fi

@@

% import the TeX file containing macro definitions
\input{tex/macro}

@@

\begin{figure}[tbp]
  \def\scale{0.38}
  \centering
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[scale=\scale]{example-instance}
    \caption{Focal point.}\label{fig:example-instance}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-lime}
      \caption{%
        Explanation by LIME\@.
        It provides contributions of each feature to the output,
        but does not explicitly indicate its scope.
      }\label{fig:example-lime}
      \vspace{0.4cm}
    \end{subfigure}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-anchor}
      \caption{%
        Explanation by Anchor.
        It provides the effective scope of the explanation,
        but does not show the influence of each feature.
      }\label{fig:example-anchor}
    \end{subfigure}
  \end{subfigure}
  \hspace{0.3cm}
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \ifnum\mode=0
      \vspace{-1.79cm}
    \else
      \vspace{-2.24cm}
    \fi
    \includegraphics[scale=\scale]{example-rlime}
    \caption{%
      Explanation by R-LIME\@.
      It provides both contributions of each feature and
      its effective scope.
    }\label{fig:example-rlime}
  \end{subfigure}
  \caption[Example of explanations by LIME, Anchor and R-LIME]{%
    Example of explanations by LIME~\cite{ribeiro2016why},
    Anchor~\cite{ribeiro2018anchors} and R-LIME (our proposed method).
    for a sentiment prediction model.
  }\label{fig:example}
\end{figure}
In this paper,
we focus on local model-agnostic methods.
An example of explanations by LIME and Anchor
for a sentiment prediction model is illustrated in \cref{fig:example}.
LIME~\cite{ribeiro2016why} approximates the complex decision boundary
locally using a linear model around the given focal point,
then provides the weights of the linear model as the contribution of each feature
to the output.
The explanation by LIME (\cref{fig:example-lime}) provides suggests that
the word ``not'' mainly contributes to the positive prediction,
but does not explicitly indicate the effective scope.
Without the scope,
users might mistakenly apply the knowledge derived from the explanation
to other instances far from the focal point,
potentially leading to misunderstanding of the black-box model's behavior
\cite{ribeiro2018anchors}.
For example,
user may apply the insights derived from the explanation
to the sentence ``This book is not good.''
and mistakenly conclude that the word ``not''
mainly contributes to the positive prediction for this sentence as well,
which is incorrect obviously.
Anchor~\cite{ribeiro2018anchors} represents a rectangular region
containing the focal point,
that maximizes the probability of the black-box classifier outputting
the same label as the focal point.
While Anchor provides an effective scope of the explanation,
the information it includes is less than LIME,
limiting its utility~\cite{ribeiro2018anchors}.
The explanation by Anchor (\cref{fig:example-anchor})
suggests that changing words other than ``not'' and ``bad''
has little impact on the classifier's output.
While it clearly cannot be applied to the sentence ``This book is not good''
because of not including the word ``bad'',
the explanation does not provide details about the influence of each word,
resulting in less user insight into the model's behavior compared to LIME.

@@

\subsection{Algorithm}\label{sec:alg}
{%
  \begin{figure}[t]
    \centering
    \includegraphics[width=0.32\textwidth]{visual-rlime1}
    \includegraphics[width=0.32\textwidth]{visual-rlime2}
    \includegraphics[width=0.32\textwidth]{visual-rlime3}
    \caption[Overview of the R-LIME algorithm]{%
      Overview of the R-LIME algorithm.
      The progression of the algorithm is illustrated from left to right.
      The solid line represents the rectangular region $A$,
      and the dashed line represents the linear approximation model $g$
      learned within $A$.
      The initial value of $A$ is an empty rule (entire input space),
      and predicates are added to $A$, reducing coverage.
      The process continues until $\Prec(A)\ge\tau$,
      at which point the rule with the maximum coverage is output.
    }
  \end{figure}
  \def\myidt{\hspace{\algorithmicindent}}
  \ifnum\mode=0
    \begin{algorithm}[t]
      \small
      \input{tex/alg1}
    \end{algorithm}
  \else
    \begin{algorithm}[p]
      \small
      \input{tex/alg1}
    \end{algorithm}
  \fi

  \foreach \alg in {2,3,4}{%
      \begin{algorithm}[p]
        \small
        \input{tex/alg\alg}
      \end{algorithm}
    }
}

@@

      \def\AB#1{\ifnum#1=0 A\else B\fi}
      \def\mylabel#1{\ifnum#1=0 \label{fig:A-instance}\else \label{fig:B-instance}\fi}
      \ifnum\mode=0
        \renewcommand{\arraystretch}{0.97}
      \else
        \renewcommand{\arraystretch}{1.02}
      \fi
      \begin{figure}[tbp]
        \foreach\a in {0,1}{%
            \centering
            \begin{subfigure}{\textwidth}
              \ifnum\mode=0
                \small
              \fi
              \centering
              \begin{tabular}{p{14em}m{16em}}
                \toprule
                \csvreader[no head, late after line= \\]{%
                  \dir/\index{\a}.csv
                }{}{%
                \ifnum\thecsvrow=16 \midrule\fi\csvcoli & \csvcolii
                }
                \bottomrule
              \end{tabular}
              \caption{Instance~\AB{\a}}\mylabel{\a}
              \vspace{15pt}
            \end{subfigure}
          }
        \vspace{-15pt}
        \caption[Two instances sampled from recidivism dataset]{%
          Two instances sampled from recidivism dataset.
        }\label{fig:instance}
      \end{figure}
    }
      \ifnum\mode=0
        \def\scale{0.32}
        \def\imgwidth{0.45\textwidth}
        \def\hspacebase{\hspace{-0.5em}}
      \else
        \def\scale{0.315}
        \def\imgwidth{0.495\textwidth}
        \def\hspacebase{\hspace{-1.5em}}
      \fi
      \def\vspacebase{\vspace{0.5em}}
      \def\vspacebeforecaption{\vspace{-0.4em}}
      \begin{figure}[p]
        \centering
        \begin{subfigure}[t]{\imgwidth}
          \hspacebase
          \hspace{1.0em}
          \includegraphics[scale=\scale]{\dir/lime-\Asample}
          \vspacebeforecaption
          \caption{LIME}\label{fig:A-lime}
          \vspacebase
        \end{subfigure}
        \begin{subfigure}[t]{\imgwidth}
          \hspacebase
          \hspace{0.5em}
          \includegraphics[scale=\scale]{\dir/newlime-\Asample-70}
          \vspacebeforecaption
          \caption{R-LIME ($\tau=0.70$)}\label{fig:A-rlime-70}
          \vspacebase
        \end{subfigure}
        \begin{subfigure}[t]{\imgwidth}
          \hspacebase
          \includegraphics[scale=\scale]{\dir/newlime-\Asample-80}
          \vspacebeforecaption
          \caption{R-LIME ($\tau=0.80$)}\label{fig:A-rlime-80}
        \end{subfigure}
        \begin{subfigure}[t]{\imgwidth}
          \hspacebase
          \hspace{-0.5em}
          \includegraphics[scale=\scale]{\dir/newlime-\Asample-90}
          \vspacebeforecaption
          \caption{R-LIME ($\tau=0.90$)}\label{fig:A-rlime-90}
        \end{subfigure}
        \caption[Explanation for Instance A by LIME and R-LIME]{%
          Explanation for Instance A by LIME and R-LIME.
        }\label{fig:A}
      \end{figure}
      \begin{figure}[p]
        \centering
        \begin{subfigure}[t]{\imgwidth}
          \hspacebase
          \hspace{-0.7em}
          \includegraphics[scale=\scale]{\dir/lime-\Bsample}
          \vspacebeforecaption
          \caption{LIME}\label{fig:B-lime}
          \vspacebase
        \end{subfigure}
        \begin{subfigure}[t]{\imgwidth}
          \hspacebase
          \hspace{0.5em}
          \includegraphics[scale=\scale]{\dir/newlime-\Bsample-70}
          \vspacebeforecaption
          \caption{R-LIME ($\tau=0.70$)}\label{fig:B-rlime-70}
          \vspacebase
        \end{subfigure}
        \begin{subfigure}[t]{\imgwidth}
          \hspacebase
          \hspace{-0.8em}
          \includegraphics[scale=\scale]{\dir/newlime-\Bsample-80}
          \vspacebeforecaption
          \caption{R-LIME ($\tau=0.80$)}\label{fig:B-rlime-80}
        \end{subfigure}
        \begin{subfigure}[t]{\imgwidth}
          \hspacebase
          \hspace{-0.8em}
          \includegraphics[scale=\scale]{\dir/newlime-\Bsample-90}
          \vspacebeforecaption
          \caption{R-LIME ($\tau=0.90$)}\label{fig:B-rlime-90}
        \end{subfigure}
        \caption[Explanation for Instance B by LIME and R-LIME]{%
          Explanation for Instance B by LIME and R-LIME.
        }\label{fig:B}
      \end{figure}
    }
  }
\subsubsection{Experimental Results}
The results of the experiment are shown in \cref{fig:A,fig:B}.
The values assigned to each attribute name represent the contribution
(weights of the learned linear classifier)
to the output of the black-box classifier,
normalized such that the absolute sum is 1.
The figures also display the 5 attributes with the highest absolute contribution.

@@

\subsection{Changes in Reward Distribution in Best Arm Identification}\label{sec:reward}
{%
  \renewcommand{\arraystretch}{1.1}
  \begin{table}[tbp]
    \centering
    \caption[Deviation between the estimated accuracy and the true accuracy]{%
      Deviation between the estimated accuracy and the true accuracy.
      Deviation was relatively small considering confidence level $1-\delta=0.95$.
    }\label{tab:reward}
    \input{tex/tab_exp_adv}
  \end{table}
}
In R-LIME,
the problem of selecting the rule with the highest accuracy is formulated
as the optimal arm identification problem in multi-armed bandit theory,
solved using the KL-LUCB algorithm \cite{kaufmann2013information}.
However, this algorithm assumes that the reward distribution remains constant,
while in R-LIME,
the reward distribution (accuracy of the linear approximation)
changes with every update of the approximation model after sampling.
Therefore, rewards obtained at an early stage
might influence the estimated value and deviate from the true value.

@@

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\ifnum\mode=0
  \bibliographystyle{style/prml}
\else
  \bibliographystyle{style/splncs04}
\fi
\bibliography{MyRef/ref}
%
\end{document}
